Cluster Difference Imaging Photometric Survey. III. An Open Cluster Spanning 500\,Parsecs

### 2021-04-05 cluster_paper/ms.tex
update ms

        \vspace{-0.5cm}
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f3c.pdf}
			\includegraphics[width=0.49\textwidth]{f3d.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf TESS rotations periods and dereddened Gaia colors for \cn.}
    {\it Top row}: \cn\ split into its core and halo.  Sets
    $\mathcal{A}$ ({\it left columns}) and $\mathcal{B}$ ({\it right
    columns}) undergo successive degrees of automated cleaning,
    described in Section~\ref{subsubsec:cluster}.  The Pleiades
    \citep[120\,Myr;][]{rebull_rotation_2016a} and Praesepe
    \citep[650\,Myr;][]{douglas_poking_2017} are shown for reference.
    Stars in the core and the halo of NGC\,2516 appear
    gyrochronologically coeval with the Pleiades at $(Bp-Rp)_0$$<$1.2.
    There is a narrow range of color, from $1.2<(Bp-Rp)_0<1.7$, in
    which some NGC\,2516 members appear slightly elevated with respect
    to the Pleiades, suggesting an older age.  
    Stars redder than $(Bp-Rp)_0\gtrsim 2.2$ (later than M2V) 
    are fainter than the magnitude limit of our survey.
    {\it Bottom row}: \cn\
    colored by binarity indicators.  Sources with $\mathrm{RUWE}>1.2$
    appear in red (astrometric binaries; 11\% of stars).  Overplotted
    in orange are sources at least 0.3\,mag above an empirically fit
    cluster isochrone (photometric binaries; 20\% of stars).
		\label{fig:rot}
	}
\end{figure*}

@@

The bottom row of Figure~\ref{fig:rot} shows the same data, with the
points colored with indicators of binarity.  To assess astrometric
binarity, we used the renormalized unit weight error (RUWE) from Gaia
EDR3.  For plotting purposes, we labeled anything with RUWE exceeding
1.2 as an astrometric binary (11\% of the overall cluster sample; see
{\it e.g.}, \citealt{belokurov_unresolved_2020}).  To assess
photometric binarity, we fitted a spline to Figure~\ref{fig:hr} by
hand, and labelled any points brighter than 0.3\,mag above the cluster
sequence as photometric binaries.  This included 20\% of the overall
cluster sample.  Many of the resulting candidate binaries, though not
all of them, appear underneath the ``slow sequence''.  We speculate on
possible explanations in Section~\ref{disc:rotn}.

@@

\subsubsection{The slow and fast sequences}
The existence of a slow sequence is the main premise of gyrochronology
\citep[{\it e.g.},][]{barnes_rotational_2003}.  While the initial
spindown relation ($P_{\rm rot}\sim t^{1/2}$) proposed by
\citet{skumanich_time_1972} has not worked in detail, the fact that it
is simple and can be understood to first order through magnetic
breaking \citep{weber_angular_1967} has led to more systematic efforts
at age-dating stars through their rotation periods.

The other features of the rotation-color diagram are in some sense
affronts to the magnetic breaking model, and the general question of
what produces the rapid rotators has not yet been definitively
answered.  Models that introduces variable core-envelope decoupling
timescales can reproduce each type of population
\citep{Irwin_NGC2516_2007}, but the question then shifts to why these
decoupling timescales should be differ for stars that are otherwise
identical.  Another possible explanation invokes a spontaneous and
random change of magnetic field topology that cause the dynamo to
couple more strongly with the stellar wind
\citep{brown_metastable_2014}.  Other processes could be external the
star entirely: \citet{qureshi_signature_2018} for instance have
reported that mergers between giant planets and the star could explain
a non-neglible fraction of the rapid rotators.

A separate hypothesis that we can test from the existing data is the
idea that 
the rapid rotators are somehow linked to binarity.
Analyses by {\it e.g.}, \citet{gillen_ngts_2020} and
\citet{simonian_rapid_2019} have reported that
rapid rotators in young clusters and in the field are correlated
with binarity.
Related effects could include tidal synchronization for the shortest
period binaries,
or some form of disk truncation that when combined with magnetic locking
(CITE)
could produce the rotation period anomaly.
The lower panels of Figure~\ref{fig:rot} show our best attempt to
identify unresolved binaries in the \cn\ rotation sample.
Binaries resolved by Gaia were already excluded in our initial
definitions of the samples.
WHAT IS THE FRACTION OF FAST ROTATORS THAT ARE BINARY?
WHAT CONCLUSION CAN BE MADE?
%FIXME CALCULATE!!!!
%FIXME CALCULATE!!!!
%FIXME CALCULATE!!!!
%FIXME CALCULATE!!!!

\subsubsection{The slowest K dwarfs of NGC\,2516}
%FIXME need to rewrite this paragraph? (Or maybe not, because it's
%already in the absolute age section...
A few conclusions can be drawn by comparing the \cn\ rotation periods
with those of the Pleiades and Praesepe.  The first is that in both
stages of cleaning, the core and halo of \cn\ each show
gyrochronological ``slow sequences''.  These sequences appear to be
coeval with each other, and at first order appear to overlap
with the Pleiades sequence.

  \paragraph{Stalling}
  Curtis19 found NGC6811 was stalling in the K dwarfs, i.e., Skumanich
  spindown is not monotonic in time.
  We're seeing latest K dwarfs spin slowest, earlier FGKs don't show
  much change.

>>>>

        \vspace{-0.5cm}
		\subfloat{
			\includegraphics[width=0.49\textwidth]{f3c.pdf}
			\includegraphics[width=0.49\textwidth]{f3d.pdf}
		}
	\end{center}
	\vspace{-0.7cm}
  \caption{ {\bf TESS rotations periods and dereddened Gaia colors for \cn.}
    {\it Top row}: \cn\ split into its core and halo.  Sets
    $\mathcal{A}$ ({\it left columns}) and $\mathcal{B}$ ({\it right
    columns}) undergo successive degrees of automated cleaning,
    described in Section~\ref{subsubsec:cluster}.  The Pleiades
    \citep[120\,Myr;][]{rebull_rotation_2016a} and Praesepe
    \citep[650\,Myr;][]{douglas_poking_2017} are shown.
    Stars in the core and the halo of NGC\,2516 appear
    gyrochronologically coeval with the Pleiades at $(Bp-Rp)_0$$<$1.2.
    There is a narrow range of color, from $1.2<(Bp-Rp)_0<1.7$, in
    which some NGC\,2516 members appear slightly elevated with respect
    to the Pleiades, suggesting an older age.  
    Stars redder than $(Bp-Rp)_0\gtrsim 2.2$ (later than M2V) 
    are fainter than the magnitude limit of our survey.
    {\it Bottom row}: \cn\
    colored by binarity indicators.  Sources with $\mathrm{RUWE}>1.2$
    appear in red (astrometric binaries; 11\% of stars).  Overplotted
    in orange are sources at least 0.3\,mag above an empirically fit
    cluster isochrone (photometric binaries; 20\% of stars).
		\label{fig:rot}
	}
\end{figure*}

@@

The bottom row of Figure~\ref{fig:rot} shows the same data, with the
points colored with indicators of binarity.  To assess astrometric
binarity, we used the renormalized unit weight error (RUWE) from Gaia
EDR3.  For plotting purposes, we labeled anything with RUWE exceeding
1.2 as an astrometric binary (11\% of the overall cluster sample; see
{\it e.g.}, \citealt{belokurov_unresolved_2020}).  To assess
photometric binarity, we fitted a spline to Figure~\ref{fig:hr} by
hand, and labelled any points brighter than 0.3\,mag above the cluster
sequence as photometric binaries.  This included 20\% of the overall
cluster sample.  Many of the resulting candidate binaries, though not
all of them, appear underneath the slow rotation sequence.  In set
$\mathcal{B}$, over the color range where the slow and fast sequence
are present ($0.5<(Bp-Rp)_0<1.3$), there are 281 stars, of which 103
(37\%) show signs of binarity.  Dividing these 281 stars into ``fast''
and ``slow'' sequences by eye,  the fraction of stars showing signs of
binarity in the slow and fast rotation subsamples are 30\% (63/210)
and 56\% (40/71), respectively.  We speculate on possible explanations
in Section~\ref{disc:rotn}.

@@

\subsubsection{The slow and fast sequences}
The existence of a slow sequence is the main premise of gyrochronology
\citep[{\it e.g.},][]{barnes_rotational_2003}.  While the initial
spindown relation ($P_{\rm rot}\sim t^{1/2}$) proposed by
\citet{skumanich_time_1972} has not worked in detail, the 
general idea is correct, and the first-order effect can be understood through magnetic
breaking \citep{weber_angular_1967}.
%More systematic efforts at age-dating stars through their rotation periods.

The other features of the rotation-color diagram are in some sense
affronts to the magnetic breaking model, and the general question of
what produces the rapid rotators has not yet been definitively
answered.  Models that introduce variable core-envelope decoupling
timescales can reproduce each type of population
\citep{Irwin_NGC2516_2007}, but the question then shifts to why these
decoupling timescales should be differ for stars that are otherwise
identical.  Another possible explanation invokes a spontaneous and
random change of magnetic field topology that causes the dynamo to
couple more strongly with the stellar wind
\citep{brown_metastable_2014}.  Other processes could be external the
star entirely: \citet{qureshi_signature_2018} for instance have
reported that mergers between giant planets and the star could explain
a non-neglible fraction of the rapid rotators.

A separate hypothesis that we can test using the existing data is the
idea that rapid rotation can be explained through binarity.  Analyses
by {\it e.g.}, \citet{gillen_ngts_2020} and
\citet{simonian_rapid_2019} have reported that rapid
rotation in young clusters and in the field is correlated with
binarity.  Related effects could include tidal synchronization for the
shortest period binaries, or alternatively disk truncation that when
combined with magnetic locking {\bf (CITE)} could produce the rotation
period bimodality.

The lower panels of Figure~\ref{fig:rot} show our attempt at
identifying unresolved binaries in the \cn\ rotation sample.
Binaries resolved by Gaia were already excluded in our initial
definitions of the samples.
In \cn\, we see that
the fraction of stars showing signs of binarity in the slow and fast
rotation subsamples are 30\% (63/210) and 56\% (40/71); the fast
rotators show a preference for binarity.
This correlation is in line with the earlier findings noted above,
though we emphasize many rapid rotators in \cn\ (34\%) have no 
evidence for being binary.

Comparing the explanations of disk-locking against tidal
synchronization, the population statistics do seem to strain the
possibility of tidal synchronization.  A quarter of the \cn\ members
are rapidly rotating.  In comparison, in the field half of Sun-like
stars are binaries, and $\approx$9\% have periods below 100\,days
\citep{raghavan_survey_2010}.  If we assume  that all
binaries with sub-100\,day periods become tidally synchronized, we
can only explain a rapid rotator occurrence rate of
$\approx$5\%.

A separate effect is that on the slow rotation sequence, 
Figure~\ref{fig:rot}} shows that the
binary stars appear to be either preferentially redder, or to have
faster rotation periods than the single stars.  One likely
explanation for this could be that the unresolved binaries have a
component contributing additional red light to the system, skewing
the color measurement of the primary.  Whether any physical
effects could be at play remains a question for future
work.

% \subsubsection{The slowest K dwarfs of NGC\,2516}
% %FIXME need to rewrite this paragraph? (Or maybe not, because it's
% %already in the absolute age section...
% A few conclusions can be drawn by comparing the \cn\ rotation periods
% with those of the Pleiades and Praesepe.  The first is that in both
% stages of cleaning, the core and halo of \cn\ each show
% gyrochronological ``slow sequences''.  These sequences appear to be
% coeval with each other, and at first order appear to overlap
% with the Pleiades sequence.
% 
%   \paragraph{Stalling}
%   Curtis19 found NGC6811 was stalling in the K dwarfs, i.e., Skumanich
%   spindown is not monotonic in time.
%   We're seeing latest K dwarfs spin slowest, earlier FGKs don't show
%   much change.
